E-Graphs With Bindings

### 2025-01-27 LICS_2025/e_graphs_with_bindings/sections/soundness.tex
clean-up

\begin{definition}
\label{def:rewrite}
    We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo SMC laws via a rewrite rule $\langle l, r \rangle$ (notation, $\leadsto_{\langle l, r \rangle}$) if they are representable as
    $
    f = c_{1};(id_{k} \otimes l);c_{2} $ and $ g = c_{1};(id_{k} \otimes r);c_{2}
    $.
\end{definition}
This notion generalises to rewriting of closed $\Sigma$-terms
\begin{definition}
    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
    $
    f = c_{1};(id_{k} \otimes d);c_{2} $ and $ g = c_{1};(id_{k} \otimes e);c_{2}
    $
    and $d \leadsto_{\langle l, r \rangle} e$
\end{definition}
\begin{lemma}
\label{lemma:normal_form}
    Every closed $\Sigma^{+}$ term $f$ can be equivalently represented as
    $
    f_{1} + \ldots + f_{n}
    $ modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
\end{lemma}
% \begin{proof}
% Such forms are given by orienting $\catname{SLat}$-equations such that $+$ gets propagated to the top, as, for example $f \otimes (g + h) = f \otimes g + g \otimes h$.
% Most notably, we can also escape lambda abstractions by using distributivity law~\ref{law:distributivity}.
% \end{proof}
This takes us to the following definition of rewriting for closed $\Sigma^{+}$-terms.
\begin{definition}
    For closed $\Sigma^{+}$-term $f,g$, $f \leadsto_{\langle l, r \rangle}$ SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
    $
    f = f_{1} + \ldots + f_{i} + \ldots + f_{n} $ and $ g = f_{1} + \ldots + f_{j} + \ldots + f_{n}
    $
    such that there is a permutation $\sigma$ such that there exist indices $i,j$ and $\sigma(f_{i}) \leadsto f_{j}$.
\end{definition}
Next, to formulate the correspondence between rewriting systems we first define interpretation $\llbracket - \rrbracket$ of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Figure~\ref{fig:ev_and_lambda}.

@@

Then we can make $\WellTypedMdaEcospans$ into $\catname{SLat}$-category by defining a join of two cospans and introducing DPOI rewrite rules to satisfy the laws of $\catname{SLat}$ symmetric monoidal category.
We also add a structural rewrite schema rule corresponding to distributivity law~\ref{law:distributivity}.
The join of two cospans is defined as in Figure~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
For example, the distributivity rule for $\otimes$ is given by the following schema rule
\[
\adjustbox{width=\linewidth}{
\tikzfig{./figures/semilattice_rule_1}~.
}
\]

@@

\begin{proposition}[Proposition 26 and Proposition 27~\cite{fscd}]
\label{prop:fscd}
For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{blue}{E}$.
\end{proposition}

>>>>

\begin{definition}
\label{def:rewrite}
    We say that a $\Sigma$-term $f$ rewrites to a $\Sigma$-term $g$ modulo SMC laws via a rewrite rule $\langle l, r \rangle$ (notation, $\leadsto_{\langle l, r \rangle}$) if they are representable as
    $
    f = c_{1};(id_{k} \otimes l);c_{2} $ and $ g = c_{1};(id_{k} \otimes r);c_{2}
    $.
\end{definition}
This notion generalises to rewriting of closed $\Sigma$-terms
\begin{definition}
    For closed-$\Sigma$-terms $f,g$, $f \leadsto_{\langle l, r \rangle}$ modulo SMC laws if they are representable as either in Definition~\ref{def:rewrite}, or as 
    $
    f = c_{1};(id_{k} \otimes d);c_{2} $ and $ g = c_{1};(id_{k} \otimes e);c_{2}
    $
    and $d \leadsto_{\langle l, r \rangle} e$
\end{definition}
\begin{lemma}
\label{lemma:normal_form}
    Every closed $\Sigma^{+}$ term $f$ can be equivalently represented as
    $
    f_{1} + \ldots + f_{n}
    $ modulo SMC laws and $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} where none of $f_{i}$ contain the join operator.
\end{lemma}
% \begin{proof}
% Such forms are given by orienting $\catname{SLat}$-equations such that $+$ gets propagated to the top, as, for example $f \otimes (g + h) = f \otimes g + g \otimes h$.
% Most notably, we can also escape lambda abstractions by using distributivity law~\ref{law:distributivity}.
% \end{proof}
This takes us to the following definition of rewriting for closed $\Sigma^{+}$-terms.
\begin{definition}
    For closed $\Sigma^{+}$-term $f,g$, $f \leadsto_{\langle l, r \rangle}$ SMC laws, $\catname{SLat}$-equations and distributivity law~\ref{law:distributivity} if they are representable as
    $
    f = f_{1} + \ldots + f_{i} + \ldots + f_{n} $ and $ g = f_{1} + \ldots + f_{j} + \ldots + f_{n}
    $
    such that there is a permutation $\sigma$ such that there exist indices $i,j$ and $\sigma(f_{i}) \leadsto f_{j}$.
\end{definition}
Next, to formulate the correspondence between rewriting systems we first define interpretation $\llbracket - \rrbracket$ of closed $\Sigma^{+}$ terms in $\WellTypedMdaEcospans$.
It follows by first defining the interpretation of operators by extending the interpretation of plain $\Sigma$ terms as morphisms in $\MdaCospans$ as given in Appendix~\ref{sec:appendix:interpretation}.
The only missing cases, $\llbracket \textsf{ev}_{A,B} \rrbracket$ and $\llbracket \Lambda_{A,B,C} \rrbracket$, are given in Fig.~\ref{fig:ev_and_lambda}.

@@

Then we can make $\WellTypedMdaEcospans$ into $\catname{SLat}$-category by defining a join of two cospans and introducing DPOI rewrite rules to satisfy the laws of $\catname{SLat}$ symmetric monoidal category.
We also add a structural rewrite schema rule corresponding to distributivity law~\ref{law:distributivity}.
The join of two cospans is defined as in Fig.~\ref{fig:f+g} and the quotienting is done through DPOI rewrite schema rules, one for each axiom.
For example, the distributivity rule for $\otimes$ is given by the following schema rule
\[
\adjustbox{width=\linewidth}{
\tikzfig{./figures/semilattice_rule_1}~.
}
\]

@@

\begin{proposition}[Proposition 26 and Proposition 27~\cite{fscd}]
\label{prop:fscd}
For two closed $\Sigma$-terms $f$ and $g$ and a $\mathcal{E}$ equation $l = r$  $f \leadsto_{\langle l, r \rangle} g$ if and only if $\llbracket f \rrbracket \Rrightarrow_{\langle \llbracket l \rrbracket, \llbracket r \rrbracket \rangle} \llbracket g \rrbracket$ where the latter is defined in a subcategory of $\WellTypedMdaEcospans$ where carriers only contain edges from $\colorbox{yellow}{E} \cup \textcolor{closed-color}{E}$.
\end{proposition}
